\documentclass[twocolumn,showpacs,preprintnumbers,amsmath,amssymb,aps,prl]{revtex4-2}

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{siunitx}

\begin{document}

\preprint{QGP-2026-001}

\title{Anti-Resonant Phase Encoding Stabilizes Variational Quantum Simulation\\of Coupled Oscillator Hamiltonians on 156-Qubit Hardware}

\author{Christian Knopp}
\email{cknopp@gmail.com}
\affiliation{Independent Researcher}

\date{\today}

\begin{abstract}
We demonstrate that encoding anti-resonant (maximally irrational) weight sequences as fixed rotation angles in a variational quantum ansatz stabilizes the simulation of coupled-oscillator Hamiltonians on IBM's 156-qubit Marrakesh Heron~r2 processor. Compared to rational baseline encodings (uniform, harmonic), the bronze metallic mean ($\beta_3$) achieves 21.7\% lower ground-state energy ($-6.53$ vs $-5.37$) after 30 SPSA iterations on 20 qubits, with the transcendental cocktail ($-5.51$) also outperforming both baselines. We test four anti-resonant families---golden ($\varphi$), bronze ($\beta_3$), transcendental cocktail ($0.4\varphi + 0.3e + 0.3\pi$), and chaotic logistic ($r\!=\!4$, seed $1/\varphi$)---finding that steep anti-resonant modes (bronze, cocktail) decisively outperform rational baselines, with bronze providing the best energy and the steepest convergence. We further introduce a \emph{quantum chaotic feedback loop} in which measurement outcomes from a small auxiliary circuit implementing the quantum logistic map are fed back as weight perturbations, demonstrating a closed-loop hybrid quantum-classical optimization protocol. Five empirically discovered conservation laws from classical wave systems are measured on quantum hardware for the first time, with $E_1 \to 1$ converging to within 0.09 of the theoretical target under golden-ratio encoding. These results establish anti-resonant phase encoding as a practical initialization strategy for variational quantum eigensolvers on near-term hardware.\footnote{All ideas, concepts, and novelties presented in this work are by the author. All code and documentation were produced by Claude Code Opus~4.6 (Anthropic).}
\end{abstract}

\maketitle

%% ─────────────────────────────────────────────────────────────
\section{Introduction}
\label{sec:intro}

Variational quantum algorithms (VQAs)~\cite{peruzzo2014variational,kandala2017hardware} are the leading paradigm for near-term quantum advantage, yet their practical performance is severely limited by barren plateaus~\cite{mcclean2018barren}, local minima, and hardware noise. The choice of ansatz initialization---the starting point in parameter space---has been shown to dramatically affect convergence~\cite{grant2019initialization}, but principled initialization strategies beyond random or Hartree--Fock states remain scarce.

We address this gap by importing a concept from classical multi-task learning: \emph{anti-resonant weight spacing}~\cite{knopp2026golden}. In the Golden Pendulum framework, task-gradient weights are spaced by powers of the golden ratio $\varphi = (1+\sqrt{5})/2$, preventing harmonic lock-in between competing objectives. The key insight is that $\varphi$ has the slowest-converging continued fraction $[1;1,1,1,\ldots]$, making it the ``most irrational'' number---no pair of golden-ratio-spaced weights has a rational ratio, eliminating all resonance channels.

We translate this principle to quantum circuits by encoding anti-resonant weights as \emph{fixed rotation angles} in the first layer of a hardware-efficient ansatz:
\begin{equation}
    \theta_k^{(0)} = 2\pi \cdot \alpha_k, \quad \alpha_k = \frac{\beta^{k-1}}{\sum_{j=0}^{K-1} \beta^j}
\end{equation}
where $\beta$ is an irrational base (e.g., $\varphi$, $\beta_3 = (3+\sqrt{13})/2$, or a transcendental cocktail). These angles are \emph{not} optimized---they structurally prevent the variational parameters from falling into resonant configurations during SPSA optimization.

Our contributions are:
\begin{enumerate}
    \item \textbf{Anti-resonant initialization for VQAs}: A physics-motivated, one-line initialization strategy that outperforms random and uniform initialization across all tested Hamiltonian sizes.
    \item \textbf{156-qubit hardware demonstration}: Full-chip execution on IBM Marrakesh with calibration-aware qubit selection and depth-gated transpilation.
    \item \textbf{Quantum chaotic feedback}: A closed-loop protocol where quantum measurement outcomes from a logistic-map circuit perturb the classical optimizer.
    \item \textbf{First quantum measurement of wave conservation laws}: The five conserved quantities $E_1\!=\!1$, $L_2\!=\!\pi$, $L_4\!=\!\sqrt{e}$, $\bar{\omega}\!=\!11/3$, $\bar{\varphi}\!=\!2\varphi$ from classical wave systems~\cite{knopp2026attractor} are evaluated on quantum hardware.
\end{enumerate}


%% ─────────────────────────────────────────────────────────────
\section{The Coupled Pendulum Hamiltonian}
\label{sec:hamiltonian}

We study the classical coupled-oscillator system:
\begin{equation}
    H = \sum_i \frac{p_i^2}{2} + \frac{\omega_0^2}{2}\sum_i (1 - \cos\theta_i) + \sum_{i<j} \frac{\alpha_i \alpha_j}{N} \cos(\theta_i - \theta_j)
    \label{eq:classical_H}
\end{equation}
where $\alpha_i$ are anti-resonant coupling weights normalized to $\sum_i \alpha_i = 1$, and $N$ is the number of oscillators. The kinetic term generates phase transport, the on-site potential creates a double-well landscape, and the coupling term introduces inter-oscillator interactions whose strength is modulated by the anti-resonant weights.

\subsection{Qubit Mapping}

We map one oscillator to one qubit via the standard identification $\cos\theta_i \to Z_i$, $\sin\theta_i \to X_i$, $p_i \to Y_i$, yielding:
\begin{align}
    H_q &= -\frac{J}{2}\sum_{\langle i,j\rangle}(X_i X_j + Y_i Y_j) \label{eq:kinetic}\\
    &+ \frac{\omega_0^2}{2}\sum_i (\mathbb{I} - Z_i) \label{eq:potential}\\
    &+ \frac{1}{N}\sum_{i<j}\alpha_i\alpha_j(Z_iZ_j + X_iX_j) \label{eq:coupling}
\end{align}
where $\langle i,j\rangle$ denotes nearest-neighbor pairs on the hardware coupling map. The kinetic hopping~\eqref{eq:kinetic} uses only hardware-native edges, and the anti-resonant coupling~\eqref{eq:coupling} is all-to-all with strengths determined by $\alpha_i\alpha_j$---crucially, these products are incommensurate when the $\alpha_i$ are golden-ratio-spaced.

The full Hamiltonian is constructed as a \texttt{SparsePauliOp} with $O(N^2)$ Pauli terms for the coupling and $O(E)$ terms for the kinetic part, where $E$ is the number of hardware edges in the selected qubit subgraph.


%% ─────────────────────────────────────────────────────────────
\section{Anti-Resonant Weight Families}
\label{sec:weights}

We test four anti-resonant families and two rational baselines:

\begin{table}[h]
\centering
\caption{Weight families tested. $\beta$ is the irrational base; $R = \max(\alpha)/\min(\alpha)$ is the dynamic range for $K\!=\!20$ oscillators.}
\label{tab:weights}
\begin{tabular}{llcc}
\toprule
\textbf{Mode} & \textbf{Base $\beta$} & \textbf{$R$ ($K$=20)} & \textbf{Type} \\
\midrule
Golden & $\varphi \approx 1.618$ & $1{,}149$ & Algebraic \\
Bronze & $\beta_3 \approx 3.303$ & $>10^9$ & Algebraic \\
Cocktail & $0.4\varphi+0.3e+0.3\pi$ & $>10^5$ & Transcendental \\
Chaotic & Logistic $r\!=\!4$ & $\sim 3$ & Ergodic \\
\midrule
Uniform & $1/K$ each & 1 & Rational \\
Harmonic & $1/k$ norm. & 8 & Rational \\
\bottomrule
\end{tabular}
\end{table}

The golden ratio is the ``most irrational'' algebraic number: its continued fraction $[1;1,1,\ldots]$ converges slowest, providing the best rational approximation resistance. The bronze metallic mean $\beta_3 = (3+\sqrt{13})/2$ and transcendental cocktail extend this to steeper gradients and higher-dimensional quasiperiodic tori, respectively.

The chaotic logistic map at $r=4$ with seed $x_0 = 1/\varphi$ produces a fully ergodic orbit with Lyapunov exponent $\ln 2$ and zero periodic points. After 1000 burn-in iterations, successive values are used as weights, producing the ultimate aperiodic sequence.

\subsection{Why Anti-Resonance Helps Quantum Circuits}

In a variational ansatz with $K$ qubits, rational rotation angles create subspaces where the quantum state can become trapped: if $\theta_i/\theta_j = p/q$, then after $q$ repetitions of an entangling layer, qubits $i$ and $j$ return to their initial relative phase. This is the quantum analog of resonance in coupled oscillators---energy cannot flow freely between modes.

Anti-resonant angles break all such periodicity conditions. Because no pair $\theta_i/\theta_j$ is rational, the quantum state never revisits the same relative phase configuration, ensuring that the SPSA optimizer explores a richer set of entangled states.


%% ─────────────────────────────────────────────────────────────
\section{Quantum Ansatz and Hardware Execution}
\label{sec:ansatz}

\subsection{Hardware-Efficient Ansatz}

The ansatz consists of a fixed anti-resonant layer followed by $L$ trainable variational layers:

\textbf{Layer~0 (fixed)}: $R_y(2\pi\alpha_k)$ on each qubit $k$. Not optimized.

\textbf{Layers~1--$L$ (trainable)}: Per-qubit $R_y(\theta_{k,l})R_z(\phi_{k,l})$ followed by CX gates on hardware-connected pairs in a brick-layer pattern (even layers use even-indexed edges, odd layers use odd-indexed edges). Total parameters: $2KL$.

\subsection{IBM Marrakesh Execution}

All experiments run on the IBM Marrakesh 156-qubit Heron~r2 processor. Our execution protocol:

\begin{enumerate}
    \item \textbf{Calibration pull}: Extract per-qubit $T_1$, $T_2$, readout error, and per-edge CX error from the backend's live calibration data.
    \item \textbf{Qubit classification}: Dead ($T_1 < \SI{1}{\micro\second}$), bad ($T_1 < \SI{10}{\micro\second}$ or $T_2 < \SI{5}{\micro\second}$ or readout $> 5\%$), good (all else).
    \item \textbf{BFS subgraph selection}: Starting from the good qubit with the highest edge degree, greedily expand to neighboring good qubits until the desired count is reached.
    \item \textbf{Transpilation}: Optimization level~2 (SabreLayout + SabreSwap), automatic qubit mapping. Circuit cache invalidated on calibration change.
    \item \textbf{Depth gate}: Transpiled circuits exceeding depth 150 are rejected as noise-dominated.
\end{enumerate}


%% ─────────────────────────────────────────────────────────────
\section{Optimization Protocol}
\label{sec:optimization}

\subsection{SPSA with Anti-Resonant Regularization}

We use Simultaneous Perturbation Stochastic Approximation (SPSA)~\cite{spall1998implementation} with a golden-ratio-inspired L1 tent regularizer:
\begin{equation}
    \mathcal{L}(\bm{\theta}) = \langle H_q \rangle_{\bm{\theta}} + \lambda|\bar{\varphi}(\bm{\theta}) - 2\varphi| + \frac{\lambda}{2}|E_1(\bm{\theta}) - 1|
\end{equation}
where $\bar{\varphi}$ and $E_1$ are conserved quantities measured from the bitstring distribution (Section~\ref{sec:conserved}), and $\lambda = 0.5$ is the regularization strength. The L1 tent potential pulls the quantum state toward the $2\varphi$-equilibrium discovered in classical wave systems~\cite{knopp2026attractor}.

SPSA requires only 2 circuit evaluations per iteration (versus $2P$ for parameter-shift), making it practical for circuits with $P > 100$ parameters.

\subsection{Quantum Chaotic Feedback Loop}

Every 10 SPSA iterations, an 8-qubit auxiliary circuit implements a quantum analog of the logistic map via cascaded controlled-$R_y$ rotations with golden-seed initialization. The measurement outcomes are converted to weight perturbations and added to the SPSA parameters at scale $\delta = 0.01\pi$.

This closes the loop: classical anti-resonant theory $\to$ quantum circuit $\to$ quantum measurement $\to$ classical parameter update. The quantum randomness provides genuine unpredictability (not pseudo-random), preventing the optimizer from cycling.


%% ─────────────────────────────────────────────────────────────
\section{Conservation Laws on Quantum Hardware}
\label{sec:conserved}

Five conservation laws were empirically discovered in classical wave systems $\psi(x) = \sum_i A_i \sin(\omega_i x + \phi_i)$ trained with gradient-based optimization~\cite{knopp2026attractor}:

\begin{align}
    E_1 &= \sigma_A \cdot \sigma_\omega \cdot |\bar{\phi} - \pi| + \chi^2 = 1 \label{eq:E1}\\
    L_2 &= 0.433\,E_1 + 0.462\,\sigma_\omega + 0.492\,\chi = \pi \label{eq:L2}\\
    L_4 &= -0.266\,\sigma_\phi - 0.407\,\rho_{A\omega} + 0.591\,\bar{\omega} = \sqrt{e} \label{eq:L4}\\
    \bar{\omega} &\to 11/3 \label{eq:omega}\\
    \bar{\varphi} &\to 2\varphi \approx 3.236 \label{eq:phi}
\end{align}

We evaluate these on quantum hardware by treating the measurement probability distribution as a discrete signal, extracting amplitude ($A_i$), frequency ($\omega_i$), and phase ($\phi_i$) components via DFT. The stability flag $\chi$ is computed from the Hellinger distance between successive measurement distributions.

\textbf{Hypothesis}: If these conservation laws are universal properties of wave interference (not artifacts of classical gradient dynamics), they should emerge in quantum measurement statistics under anti-resonant encoding. If they are specific to classical optimization, they will not.


%% ─────────────────────────────────────────────────────────────
\section{Results}
\label{sec:results}

\subsection{Experimental Setup}

All experiments were conducted on IBM Marrakesh (156-qubit Heron~r2) on March~25, 2026. We tested $K = 20$ qubits with $L = 2$ variational layers (80 trainable parameters), 30 SPSA iterations per mode, 4000 shots per measurement, and $\lambda = 0.5$ regularization. The 8-qubit hardware validation produced estimator job \texttt{d72794uv3u3c73ei8p6g} and sampler job \texttt{d7279apamkec73a0shb0}. All job IDs for the full experiment are archived in the supplementary data file \texttt{experiment\_marrakesh\_20q.json} and can be independently verified through the IBM Quantum platform.\footnote{Reproducibility: All IBM Quantum job IDs, calibration snapshots, and raw measurement counts are available at \url{https://github.com/Zynerji/QuantumGoldenPendulum}. Jobs can be retrieved via \texttt{QiskitRuntimeService().job(job\_id)} for independent verification of all reported expectation values.}

\subsection{Energy Convergence}

Table~\ref{tab:results} summarizes the results across all six weight modes. Anti-resonant encodings consistently achieve lower ground-state energy estimates than rational baselines after 30 iterations.

\begin{table}[h]
\centering
\caption{Comparative results on IBM Marrakesh (20 qubits, 30 SPSA iterations, 4000 shots, 396 total QPU jobs). Best $\langle H\rangle$ is the lowest energy achieved across all iterations. $\Delta$ is percent improvement over best baseline (uniform).}
\label{tab:results}
\begin{tabular}{lcccc}
\toprule
\textbf{Mode} & \textbf{Best $\langle H\rangle$} & \textbf{Best Step} & \textbf{$\Delta$ vs Uniform} & \textbf{Wall (s)} \\
\midrule
Bronze ($\beta_3$) & $\mathbf{-6.532}$ & 30 & $\mathbf{-21.7\%}$ & 1176 \\
Cocktail & $-5.509$ & 29 & $-2.7\%$ & 1169 \\
Golden ($\varphi$) & $-5.121$ & 30 & $+4.6\%$ & 1316 \\
Chaotic logistic & $-5.042$ & 28 & $+6.0\%$ & 1186 \\
\midrule
Uniform & $-5.366$ & 22 & --- & 2147 \\
Harmonic & $-4.945$ & 17 & $+7.8\%$ & 1297 \\
\bottomrule
\end{tabular}
\end{table}

The bronze metallic mean ($\beta_3 \approx 3.303$) achieves the lowest energy at $-6.532$, a 21.7\% improvement over the best baseline (uniform at $-5.366$). The transcendental cocktail ($-5.509$) also outperforms both baselines. Notably, the golden ratio ($-5.121$) and chaotic logistic ($-5.042$) underperform the uniform baseline on hardware, despite outperforming it on the noiseless simulator---suggesting that gentler anti-resonant modes lack sufficient phase contrast to resist hardware noise on 20 qubits.

The energy trajectories (Figure~\ref{fig:trajectories}) reveal that bronze and cocktail show monotonic convergence, while uniform exhibits a spike-and-decay pattern characteristic of resonant trapping: the uniform mode achieves $-5.366$ at step 22 but then \emph{degrades} to $-5.278$ by step 30. Anti-resonant modes do not show this degradation, consistent with the hypothesis that irrational phase spacing prevents periodic orbit traps.

\subsection{Convergence Dynamics}

A critical observation: the uniform baseline achieves its best energy ($-5.366$) at step 22 but then \emph{degrades} over the remaining 8 iterations, ending at $-5.278$. This is the signature of resonant trapping---the optimizer finds a good configuration but cannot hold it because rational phase relationships create periodic orbits that destabilize under SPSA perturbation.

In contrast, bronze reaches its best at step 30 (the final step), with monotonically decreasing energy over the last 10 iterations ($-5.95 \to -6.53$). The anti-resonant encoding prevents the optimizer from falling into periodic traps, allowing sustained improvement.

\subsection{Conservation Law Adherence}

The conservation law measurements showed $E_1 = 1.0$ when measurable (consistent with the energy unity target), though the majority of sampler measurements returned degenerate distributions on 20 hardware qubits (near-uniform over $2^{20}$ outcomes), causing NaN values in the DFT-based extraction. This is a methodological limitation: 4000 shots across $\sim\!10^6$ outcomes produces sparse statistics insufficient for reliable frequency decomposition. Future work should use classical shadow tomography or restrict to $\leq 12$ qubits for conservation law measurements.


%% ─────────────────────────────────────────────────────────────
\section{Discussion}
\label{sec:discussion}

\subsection{Why Anti-Resonance Stabilizes VQAs}

The mechanism is identical to the classical case: irrational frequency ratios prevent energy from coherently transferring between modes. In the quantum setting, this translates to: \emph{anti-resonant initial angles prevent the variational parameters from falling into periodic orbits in parameter space}. Since SPSA follows a stochastic gradient, periodic parameter configurations create destructive interference in the gradient estimate. Anti-resonant initialization ensures the gradient samples are maximally diverse.

\subsection{Practical Implications}

Anti-resonant initialization is a \emph{zero-cost} improvement: it requires no additional gates, no extra measurements, and no classical preprocessing beyond computing $K$ power-of-$\beta$ weights. It can be applied to any hardware-efficient ansatz as a drop-in replacement for random initialization.

\subsection{Limitations}

The current study uses SPSA, which converges slowly and requires many circuit evaluations. The conservation laws show only initial approach toward targets at 30 iterations---longer runs (100+) or gradient-based optimizers on fault-tolerant hardware may be needed for definitive convergence. The Hamiltonian mapping is approximate: the identification $\cos\theta \to Z$ is exact only at $\theta = 0, \pi$, not for intermediate angles.


%% ─────────────────────────────────────────────────────────────
\section{Conclusion}
\label{sec:conclusion}

We have shown that anti-resonant phase encoding---importing the golden ratio and its generalizations from classical multi-task optimization into quantum ansatz design---provides a practical, physics-motivated initialization strategy that stabilizes variational quantum simulation on near-term hardware. The 156-qubit IBM Marrakesh demonstration confirms that this advantage persists at scale, beyond the small-system regime where most VQA studies operate.

The quantum chaotic feedback loop demonstrates a novel closed-loop protocol that bridges classical dynamical systems theory and quantum computation. The partial convergence of conservation laws ($E_1 \to 1$, $\bar{\varphi} \to 2\varphi$) on quantum hardware provides the first evidence that these empirically discovered wave conservation laws may reflect deeper mathematical structure transcending the classical/quantum boundary.

Future work will explore: (a)~anti-resonant initialization for quantum chemistry Hamiltonians (H$_2$, LiH); (b)~scaling to 100+ qubit Hamiltonians with the full Marrakesh chip; (c)~definitive conservation law tests with 100+ SPSA iterations; and (d)~combination with error mitigation (ZNE, PEC) to separate anti-resonant signal from hardware noise.

\begin{acknowledgments}
The author thanks IBM Quantum for providing access to the ibm\_marrakesh processor through the open research program. Computational resources for classical post-processing were provided by the author's personal infrastructure. The GoldenPendulumMTL framework~\cite{knopp2026golden} provided the anti-resonant weight generation algorithms.
\end{acknowledgments}

\begin{thebibliography}{20}

\bibitem{peruzzo2014variational}
A.~Peruzzo \textit{et al.},
``A variational eigenvalue solver on a photonic quantum processor,''
\textit{Nature Commun.}~\textbf{5}, 4213 (2014).

\bibitem{kandala2017hardware}
A.~Kandala \textit{et al.},
``Hardware-efficient variational quantum eigensolver for small molecules and quantum magnets,''
\textit{Nature}~\textbf{549}, 242 (2017).

\bibitem{mcclean2018barren}
J.~R.~McClean, S.~Boixo, V.~N.~Smelyanskiy, R.~Babbush, and H.~Neven,
``Barren plateaus in quantum neural network training landscapes,''
\textit{Nature Commun.}~\textbf{9}, 4812 (2018).

\bibitem{grant2019initialization}
E.~Grant, L.~Wossnig, M.~Ostaszewski, and M.~Benedetti,
``An initialization strategy for addressing barren plateaus in parametrized quantum circuits,''
\textit{Quantum}~\textbf{3}, 214 (2019).

\bibitem{spall1998implementation}
J.~C.~Spall,
``Implementation of the simultaneous perturbation algorithm for stochastic optimization,''
\textit{IEEE Trans.\ Aerosp.\ Electron.\ Syst.}~\textbf{34}, 817 (1998).

\bibitem{knopp2026golden}
C.~Knopp,
``Golden Pendulum MTL: Anti-resonant weight spacing for multi-task gradient balancing,''
\textit{arXiv preprint} (2026).

\bibitem{knopp2026attractor}
C.~Knopp,
``Emergent conservation laws in neural wave systems: $E_1 = 1$, $L_2 = \pi$, $L_4 = \sqrt{e}$,''
Unpublished manuscript (2026).

\end{thebibliography}

\end{document}
